How did local health infrastructure and socio-political factors within different states and counties in the United States affect the disparities in COVID-19 outcomes, and what lessons can be learned for more targeted public health preparedness and response strategies in future pandemics?

### 2024-02-11 outputs/paper/paper.qmd
added vote share graph

ghs_data <- read_excel(here::here("inputs/data/GHS_index.xlsx"))
world_map <- st_read(here::here("inputs/data/world/World_Countries__Generalized_.shp"))
life_exp <- read.csv(here::here("inputs/data/life_table.csv"), stringsAsFactors = FALSE)
```

@@

>>>>

ghs_data <- read_excel(here::here("inputs/data/GHS_index.xlsx"))
world_map <- st_read(here::here("inputs/data/world/World_Countries__Generalized_.shp"))
life_exp <- read.csv(here::here("inputs/data/life_table.csv"), stringsAsFactors = FALSE)
vote_share <- read.csv(here::here("inputs/data/Vote_share_2020_data.csv"), stringsAsFactors = FALSE)
```

@@

```{r}
#| echo: false
#| eval: true
#| warning: false

vote_data_by_state <- vote_share %>%
  group_by(state_name) %>%
  summarize(
    total_votes = sum(total_votes),
    total_gop_votes = sum(votes_gop),
    total_dem_votes = sum(votes_dem),
    total_diff = sum(diff),
    avg_per_gop = mean(per_gop),
    avg_per_dem = mean(per_dem),
    avg_per_point_diff = mean(per_point_diff)
  )

vote_data_by_state <- vote_data_by_state |>
  mutate(gop_greater = if_else(total_gop_votes > total_dem_votes, 1, 0))

```
```{r}
#| echo: false
#| eval: true
#| warning: false

# Create plot summarizing total number of votes by top 10 republican state
sorted_states <- vote_data_by_state[order(-vote_data_by_state$total_diff), ]

# Extract the top 10 states with the biggest difference
top_10_states <- head(sorted_states, 10)
# Create a data frame for plotting
plot_data <- data.frame(
  state_name = top_10_states$state_name,
  republican_votes = top_10_states$total_gop_votes,
  democrat_votes = top_10_states$total_dem_votes
)

# Convert data to long format for ggplot
plot_data_long <- tidyr::pivot_longer(plot_data, cols = c(republican_votes, democrat_votes), names_to = "party", values_to = "votes")

# Create the stacked bar plot
ggplot(plot_data_long, aes(x = state_name, y = votes, fill = party)) +
  geom_bar(stat = "identity") +
  labs(title = "Total Number of Votes by Top 10 Republican State", x = "State", y = "Total Votes", fill = "Party") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1)) +
  scale_fill_manual(values = c("republican_votes" = "red", "democrat_votes" = "blue"),
                    labels = c("Republican Votes", "Democratic Votes"))
